---
title: "Lab 9 - HPC"
author: Tao Huang
output: tufte::tufte_html
link-citations: yes
---

# Learning goals

In this lab, you are expected to learn/put in practice the following skills:

- Evaluate whether a problem can be parallelized or not.
- Practice with the parallel package.
- Use Rscript to submit jobs
- Practice your skills with Git.

## Problem 1: Think

Give yourself a few minutes to think about what you just learned. List three
examples of problems that you believe may be solved using parallel computing,
and check for packages on the HPC CRAN task view that may be related to it.

>
--LOOCV in machine larning: 'caret','mlr'.
--high-level neural networks in machine learning: 'keras'.
--Random Forest : 'Rborist ', ' randomForestSRC','varSelRF'
--Monte Carlo simulation: 'parallel'



## Problem 2: Before you

The following functions can be written to be more efficient without using
parallel:

*1. This function generates a `n x k` dataset with all its entries distributed*
*poission with mean `lambda`.*

```{r p2-fun1, eval = FALSE}
fun1 <- function(n = 100, k = 4, lambda = 4) {      #lambda----anonymous function～
  x <- NULL
  for (i in 1:n)
    x <- rbind(x, rpois(k, lambda))
  x
} 


# faster way
fun1alt <- function(n = 100, k = 4, lambda = 4) {      #set up default value inside the function
  matrix(rpois(n*k,lambda = lambda),ncol=k)
}

fun1alt(10)       #change default value(n=100) to n=10.


install.packages("microbenchmark")
# Benchmarking
microbenchmark::microbenchmark(
  fun1(),
  fun1alt(),unit='relative'
)         #take mean as a example, we can find fun1alt() 15 times faster than fun1()
```

*2.  Find the column max (hint: Checkout the function `max.col()`).*

```{r p2-fun2, eval = FALSE}
# Data Generating Process (10 x 10,000 matrix)
set.seed(1234)
x <- matrix(rnorm(1e4), nrow=10)

# Find each column's max value
fun2 <- function(x) {
  apply(x, 2, max)      #loop
}


#faster######
fun2alt <- function(x) {
  x[cbind(
    max.col(t(x)),
    1:ncol(x)
  )]          #vector computing
}

#check 
x[,1:5]
fun2(x)[1:5]
fun2alt(x)[1:5]


# Benchmarking   
###quantitate the speed. ~~VS
microbenchmark::microbenchmark(
  fun2(x),
  fun2alt(x) , unit = 'relative'
)
```

>btw, we can use matrix to select the number i want from df.
eg. x[cbind(1:10,1:10)]


## Problem 3: Parallelize everyhing

We will now turn our attention to non-parametric 
[bootstrapping](https://en.wikipedia.org/wiki/Bootstrapping_(statistics)).
Among its many uses, non-parametric bootstrapping allow us to obtain confidence
intervals for parameter estimates without relying on parametric assumptions.

The main assumption is that we can approximate many experiments by resampling
observations from our original dataset, which reflects the population. 

*This function implements the non-parametric bootstrap:*

```{r p3-boot-fun, eval = FALSE}
my_boot <- function(dat, stat, R, ncpus = 1L) {         #ncpus=1L
  
  # Getting the random indices
  n <- nrow(dat)
  idx <- matrix(sample.int(n, n*R, TRUE), nrow=n, ncol=R)
 
  # Making the cluster using `ncpus`
  # STEP 1: Make cluster
  cl<-makePSOCKcluster(ncpus)
  
  # STEP 2: set it up(export data if need) ,idx,dat...put cl
  parallel::clusterExport(cl,varlist = c("idx","dat","stat"),envir=environment())  #wach one cluster(cl) have idex,dat...
  
  # STEP 3: THIS FUNCTION NEEDS TO BE REPLACES WITH parLapply (parallelly lappy function in each cl) , put cl inside fun
  ans <- parLapply(cl,seq_len(R), function(i) {    
    stat(dat[idx[,i], , drop=FALSE])
  })
  
  # Coercing the list into a matrix
  ans <- do.call(rbind, ans)     #do.call : rbind the reach row in kist/df-- ans...
  
  # STEP 4: stop the cluster
  stopCluster(cl)
  
  ans
  
}

my_boot
```

1. Use the previous pseudocode, and make it work with parallel. Here is just an example
for you to try:

```{r p3-test-boot, eval = FALSE}
# Bootstrap of an OLS
my_stat <- function(d) coef(lm(y ~ x, data=d))

# DATA SIM
set.seed(1)
n <- 500
R <- 1e4

x <- cbind(rnorm(n))
y <- x*5 + rnorm(n)

# Checking if we get something similar as lm
ans0 <- confint(lm(y~x))

##2 core
ans1 <- my_boot(dat = data.frame(x, y),    
                stat= my_stat, 
                R = R, 
                ncpus = 2L)         # change:  ncpus=2L == 2cpu

# You should get something like this
t(apply(ans1, 2, quantile, c(.025,.975)))    #bootstrap
##                   2.5%      97.5%
## (Intercept) -0.1372435 0.05074397
## x            4.8680977 5.04539763
ans0
##                  2.5 %     97.5 %
## (Intercept) -0.1379033 0.04797344
## x            4.8650100 5.04883353
```

2. Check whether your version actually goes faster than the non-parallel version:


```{r benchmark-problem3, eval = FALSE}
##contrast the speed  1L vs 2L.
system.time(my_boot(dat = data.frame(x, y), my_stat, R = 4000, ncpus = 1L))  
#R:the times of resample??
system.time(my_boot(dat = data.frame(x, y), my_stat, R = 4000, ncpus = 4L))
```

## Problem 4: Compile this markdown document using Rscript

Once you have saved this Rmd file, try running the following command
in your terminal:

```bash
Rscript --vanilla -e 'rmarkdown::render("[full-path-to-your-Rmd-file.Rmd]")' &
```

Where `[full-path-to-your-Rmd-file.Rmd]` should be replace with the full path to
your Rmd file... :).
