---
title: "Lab 11 - Interactive Visualization"
output: html_document
author: Tao Huang
link-citations: yes
chunk_output_type: console
---
  
  
  
```{r setup, message=FALSE, echo=FALSE, warning=FALSE}
library(data.table)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(plotly)
library(DT)
library(knitr)

opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  eval=TRUE,
  echo = TRUE,
  fig.width = 7, 
  fig.align = 'center',
  fig.asp = 0.618,
  out.width = "700px")
```

# Learning Goals

- Read in and process the COVID dataset from the New York Times GitHub repository
- Create interactive graphs of different types using `plot_ly()` and `ggplotly()` functions
- Customize the hoverinfo and other plot features
- Create a Choropleth map using `plot_geo()`
- Create an interactive table using `DataTable`

# Lab Description

We will work with the COVID data presented in lecture. Recall the dataset consists of COVID-19 cases and deaths in each US state during the course of the COVID epidemic. We will explore cases, deaths, and their population normalized values over time to identify trends.

# Steps

## I. Reading and processing the New York Times (NYT) state-level COVID-19 data

### 1. Read in the data

- Read in the COVID data with data.table:fread() from the NYT GitHub repository: "https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv"
- Read in the state population data with data.table:fread() from the repository: "https://raw.githubusercontent.com/COVID19Tracking/associated-data/master/us_census_data/us_census_2018_population_estimates_states.csv"
- Merge datasets

```{r, echo=TRUE, message=FALSE}
## data extracted from New York Times state-level data from NYT Github repository
# https://github.com/nytimes/covid-19-data
## state-level population information from us_census_data available on GitHub repository:
# https://github.com/COVID19Tracking/associated-data/tree/master/us_census_data
# load COVID state-level data from NYT
### FINISH THE CODE HERE ###
cv_states <- as.data.frame(data.table::fread("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv") )
# load state population data
### FINISH THE CODE HERE ###
state_pops <- as.data.frame(data.table::fread("https://raw.githubusercontent.com/COVID19Tracking/associated-data/master/us_census_data/us_census_2018_population_estimates_states.csv"))

state_pops$abb<-state_pops$state
state_pops$state<-state_pops$state_name
#merge them~
cv_states<-merge(state_pops,cv_states,by='state')

```

### 2. Look at the data

- Inspect the dimensions, `head`, and `tail` of the data
- Inspect the structure of each variables. Are they in the correct format?
  
```{r}

dim(cv_states)
head(cv_states)
tail(cv_states)
str(cv_states)

```

### 3. Format the data

- Make date into a date variable
- Make `state` and `abb` into a factor variable
- Order the data first by state, second by date
- Confirm the variables are now correctly formatted
- Inspect the range values for each variable. What is the date range? The range of cases and deaths?
  
```{r}
cv_states_ori <- as.data.frame(data.table::fread("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-states.csv") )
state_pops_ori <- as.data.frame(data.table::fread("https://raw.githubusercontent.com/COVID19Tracking/associated-data/master/us_census_data/us_census_2018_population_estimates_states.csv"))

cv_states$date<-as.Date(cv_states$date,format="%Y-%m-%d")     #change date format (chr----> date) ！！"%Y-%m-%d"！！

#change chr to factor...   
#state
state_list<-unique(cv_states$state)
cv_states$state<-factor(cv_states$state,levels = state_list)
#abb
abb_list<-unique(cv_states$abb)      #get the elements
cv_states$abb<-factor(cv_states$abb,levels=abb_list)     


#order the data first by state, second by date
cv_states<-cv_states[order(cv_states$state,cv_states$date),]

#check
head(cv_states)
str(cv_states)


#inspect range~
range(cv_states$date)
summary(cv_states)



```

### 4. Add `new_cases` and `new_deaths` and correct outliers

- Add variables for new cases, `new_cases`, and new deaths, `new_deaths`: 
  - Hint: `new_cases` is equal to the difference between cases on date i and date i-1, starting on date i=2

- Use `plotly` for EDA: See if there are outliers or values that don't make sense for `new_cases` and `new_deaths`. Which states and which dates have strange values?

- Correct outliers: Set negative values for `new_cases` or `new_deaths` to 0

- Recalculate `cases` and `deaths` as cumulative sum of updates `new_cases` and `new_deaths`

```{r}
# Add variables for new_cases and new_deaths:
str(cv_states)

#add new var  ---"new_cases"(cases[i]-cases[i-1])
cv_states$date[1]
cv_states$state[1]



#new cases
for (i in 1:length(state_list)){
  # by state_list to subset the date by state..
  cv_sub<-subset(cv_states,state==state_list[i])
  cv_sub$new_cases <-cv_sub$cases[1]
  for (j in 2:nrow(cv_sub)){
    cv_sub$new_cases[j]<-cv_sub$cases[j]-cv_sub$cases[j-1]
  }
  cv_states$new_cases[cv_states$state==state_list[i]] <- cv_sub$new_cases
}


##same way: new deaths
for (i in 1:length(state_list)){
  # by state_list to subset the date by state..
  cv_sub<-subset(cv_states,state==state_list[i])
  cv_sub$new_deaths <-cv_sub$deaths[1]
  for (j in 2:nrow(cv_sub)){
    cv_sub$new_deaths[j]<-cv_sub$deaths[j]-cv_sub$deaths[j-1]
  }
  cv_states$new_deaths[cv_states$state==state_list[i]] <- cv_sub$new_deaths
}

tail(cv_states)



```

```{R}
##Use `plotly` for EDA ,inspect the outlier : Set negative values for `new_cases` or `new_deaths` to 0
str(cv_states)

#plotly--interactive plot~~
#for new_cases
p1<-ggplot(cv_states,
           aes(x=date, y=new_cases, color=state)) + geom_line() + geom_point(size=.5,alpha=.5)
ggplotly(p1)   #Convert from ggplot2 to plotly.
#for new_deaths
p2<-ggplot(cv_states,
           aes(x=date, y=new_deaths, color=state)) + geom_line() + geom_point(size=.5,alpha=.5)
ggplotly(p2)

#Set negative values for `new_cases` or `new_deaths` to 0
cv_states$new_cases[cv_states$new_cases<0]=0
cv_states$new_deaths[cv_states$new_deaths<0]=0



##Recalculate `cases` and `deaths` as cumulative sum of updates `new_cases` and `new_deaths`
#
for (i in 1:length(state_list)) {
  cv_subset <- subset(cv_states, state == state_list[i])

  # add starting level for new cases and deaths
  cv_subset$cases <- cv_subset$cases[1]
  cv_subset$deaths <- cv_subset$deaths[1]

  for (j in 2:nrow(cv_subset)) {
    cv_subset$cases[j] <- cv_subset$new_cases[j] + cv_subset$cases[j-1]
    cv_subset$deaths[j] <- cv_subset$new_deaths[j] + cv_subset$deaths[j-1]
  }
  # include in main dataset
  cv_states$cases[cv_states$state==state_list[i]] <- cv_subset$cases
  cv_states$deaths[cv_states$state==state_list[i]] <- cv_subset$deaths
}





```



### 5. Add additional variables

- Add population-normalized (by 100,000) variables for each variable type (rounded to 1 decimal place). Make sure the variables you calculate are in the correct format (`numeric`). You can use the following variable names:
  - `per100k` = cases per 100,000 population
  - `newper100k`= new cases per 100,000
  - `deathsper100k` = deaths per 100,000
  - `newdeathsper100k` = new deaths per 100,000

- Add a "naive CFR" variable representing `deaths / cases` on each date for each state

- Create a dataframe representing values on the most recent date, `cv_states_today`, as done in lecture

```{r}
# add population normalized (by 100,000) counts for each variable

str(cv_states)

a<-as.numeric(format(round(cv_states$cases/(cv_states$population/100000),1),nsmall = 1))
typeof(a)
?format

cv_states$per100k <- as.numeric(format(round(cv_states$cases/(cv_states$population/100000),1),nsmall = 1))
cv_states$newper100k <- as.numeric(format(round(cv_states$new_cases/(cv_states$population/100000),1),nsmall = 1))
cv_states$deathsper100k <- as.numeric(format(round(cv_states$deaths/(cv_states$population/100000),1),nsmall = 1))
cv_states$newdeathsper100k <- as.numeric(format(round(cv_states$new_deaths/(cv_states$population/100000),1),nsmall = 1))


##Add a "naive CFR" variable representing `deaths / cases` on each date for each state
cv_states$naive_CFR <- round((cv_states$deaths*100/cv_states$cases),2)


##Create a dataframe representing values on the most recent date, `cv_states_today`, as done in lecture
cv_states_today <- filter(cv_states, date==as.Date(max(cv_states$date)))

```

## II. Interactive plots

### 6. Explore scatterplots using `plot_ly()`

- Create a scatterplot using `plot_ly()` representing `pop_density` vs. various variables (e.g. `cases`, `per100k`, `deaths`, `deathsper100k`) for each state on most recent date (`cv_states_today`)
  - Use hover to identify any outliers. 
  - Remove those outliers and replot.
- Choose one plot. For this plot:
  - Add hoverinfo specifying the state name, cases per 100k, and deaths per 100k, similarly to how we did this in the lecture notes
  - Add layout information to title the chart and the axes
  - Enable `hovermode = "compare"`

```{r}
str(cv_states)
# pop_density vs. cases
#directly use plotly~~
cv_states_today %>% 
  plot_ly(x= ~pop_density, y= ~cases, 
         type = 'scatter', mode='markers', color=~ state,
         size = ~population, sizes =c(5,70), 
         marker = list(sizemode= 'diameter', opacity=.5)
         )
#remove outlier (D.C.)
cv_states_today_re<-cv_states_today %>% filter(state!="District of Columbia")
cv_states_today_re %>% 
  plot_ly(x= ~pop_density, y= ~cases, 
         type = 'scatter', mode='markers', color=~ state,
         size = ~population, sizes =c(5,70), 
         marker = list(sizemode= 'diameter', opacity=.5)
         )

# pop_density vs. per100k
#directly use plotly~~
cv_states_today_re %>% 
  plot_ly(x= ~pop_density, y= ~per100k, 
         type = 'scatter', mode='markers', color=~ state,
         size = ~population, sizes =c(5,70), 
         marker = list(sizemode= 'diameter', opacity=.5)
         )


# pop_density vs. deaths
#directly use plotly~~
cv_states_today_re %>% 
  plot_ly(x= ~pop_density, y= ~deaths, 
         type = 'scatter', mode='markers', color=~ state,
         size = ~population, sizes =c(5,70), 
         marker = list(sizemode= 'diameter', opacity=.5)
         )


# pop_density vs. deathsper100k
#directly use plotly~~
cv_states_today_re %>% 
  plot_ly(x= ~pop_density, y= ~deathsper100k, 
         type = 'scatter', mode='markers', color=~ state,
         size = ~population, sizes =c(5,70), 
         marker = list(sizemode= 'diameter', opacity=.5)
         )


```
*district of Columbia is outlier~~*
```{R}
 #Add hoverinfo specifying the state name, cases per 100k, and deaths per 100k, similarly to how we did this in the lecture notes/ Add layout information to title the chart and the axes/ Enable `hovermode = "compare"`

cv_states_today_re %>%
  plot_ly(
    x = ~pop_density, y= ~deathsper100k,
    type ='scatter', mode= 'markers', color= ~state,
    size= ~population, sizes= c(5,70), marker=list(sizemode='diameter', opacity=.5),
    
    hoverinfo= 'text',
    text = ~paste(
      paste(state,":",sep=""),
      paste("Cases per 100k:",per100k,sep=""),
      paste("Deaths per 100k:",deathsper100k, sep=""),sep="<br>")
    ) %>%
  layout(title= "Population-normalized COVID-19 deaths (per 100k) VS. population density for US states",
             yaxis= list(title="Deaths per 100k"), xaxis=list(title="Population Density"),
             hovermode= "compare")

```





### 7. Explore scatterplot trend interactively using `ggplotly()` and `geom_smooth()`

- For `pop_density` vs. `newdeathsper100k` create a chart with the same variables using `gglot_ly()`
  - What's the `geom_*()` we need here?
  - Explore the pattern between $x$ and $y$ using `geom_smooth()`
- Explain what you see. Do you think `pop_density` is a correlate of `newdeathsper100k`?
  
```{r}
### FINISH THE CODE HERE ###
p <- ggplot(cv_states_today_re, aes(x=pop_density, y=per100k, color=state, size=population)) + geom_point() + geom_smooth() 
ggplotly(p)

p2 <- ggplot(cv_states_today_re, aes(x=pop_density, y=newdeathsper100k, color=state, size=population)) + geom_point() + geom_smooth() 
ggplotly(p2)
```
*From chart, I think pop_density` is not corrected with `newdeathsper100k`*

### 8. Multiple line chart

- Create a line chart of the `naive_CFR` for all states over time using `plot_ly()`
- Use hoverinfo to identify states that had a "first peak"
- Use the zoom and pan tools to inspect the `naive_CFR` for the states that had a "first peak" in September. How have they changed over time?
  
  - Create one more line chart, for Texas only, which shows `new_cases` and `new_deaths` together in one plot. Hint: use `add_lines()`
- Use hoverinfo to "eyeball" the approximate peak of deaths and peak of cases. What is the time delay between the peak of cases and the peak of deaths?
  
```{r}
str(cv_states)

# Line chart for naive_CFR for all states over time using `plot_ly()`
#find the peak
for (i in 1:length(state_list)){
  cv_sub<-subset(cv_states,state==state_list[i])
  cv_states$maxcfr[cv_states$state==state_list[i]] <- max(cv_sub$naive_CFR)
}

plot_ly(cv_states, x = ~date, y = ~naive_CFR, color = ~state, type = "scatter", mode = "lines",
         hoverinfo= 'text',
     text = ~paste(
      paste("The peak value of naive_CFR:",maxcfr,sep=""))
     )


# Line chart for Texas showing new_cases and new_deaths together
### FINISH THE CODE HERE ###
cv_states %>% filter(state=="Texas") %>% plot_ly(x = ~date, y = ~new_cases, type = "scatter", mode = "lines") %>% add_lines(x = ~date, y = ~new_deaths, type = "scatter", mode = "lines")


```
*- The time delay is around 2 weeks.*

### 9. Heatmaps

Create a heatmap to visualize `new_cases` for each state on each date greater than April 1st, 2020
- Start by mapping selected features in the dataframe into a matrix using the **tidyr** package function `pivot_wider()`, naming the rows and columns, as done in the lecture notes
- Use `plot_ly()` to create a heatmap out of this matrix

- Create a second heatmap in which the pattern of `new_cases` for each state over time becomes more clear by filtering to only look at dates every two weeks 

```{r}
# Map state, date, and new_cases to a matrix

#build matrix!!!(new_cases,date)  the value is new_cases..
library(tidyr)
cv_states_mat <- cv_states %>% select(state, date, new_cases) %>% filter( date > as.Date("2020-04-01"))
cv_states_mat2 <- as.data.frame(pivot_wider(cv_states_mat, names_from = state, values_from = new_cases))

rownames(cv_states_mat2)<-cv_states_mat2$date
cv_states_mat2$date <- NULL
cv_states_mat2 <- as.matrix(cv_states_mat2)

##build headmap
plot_ly(x=colnames(cv_states_mat2) , y= rownames(cv_states_mat2),
        z= ~cv_states_mat2,
        type = 'heatmap',
        showscale=T
        )



##headmap2
# Create a second heatmap after filtering to only include dates every other week
filter_dates <- seq(as.Date("2020-04-01"), as.Date("2020-10-01"), by="2 weeks")

### FINISH THE CODE HERE ### 
cv_states_mat <- cv_states %>% select(state, date, new_cases) %>% filter( date %in% filter_dates )

cv_states_mat2 <- as.data.frame(pivot_wider(cv_states_mat, names_from = state, values_from = new_cases))
rownames(cv_states_mat2) <- cv_states_mat2$date
cv_states_mat2$date <- NULL
cv_states_mat2 <- as.matrix(cv_states_mat2)

# Create a heatmap using plot_ly()
plot_ly(x=colnames(cv_states_mat2), y=rownames(cv_states_mat2),
             z=~cv_states_mat2,
             type="heatmap",
             showscale=T)

```





### 10. Map

- Create a map to visualize the `naive_CFR` by state on May 1st, 2020
- Compare with a map visualizing the `naive_CFR` by state on most recent date
- Plot the two maps side by side using `subplot()`. Make sure the shading is for the same range of values (google is your friend for this)
- Describe the difference in the pattern of the CFR.

```{r}
### For May 1 2020
# Extract the data for each state by its abbreviation
cv_cfr<-cv_states %>% filter(date=="2020-05-01") %>% select(state, abb, naive_CFR, cases, deaths)
cv_cfr$state_name <-cv_cfr$state
cv_cfr$state <-cv_cfr$abb
cv_cfr$abb<- NULL

#create hover text
cv_cfr$hover <- with(cv_cfr, paste(state_name, '<br>', "CFR: ", naive_CFR, '<br>', "Cases: ", cases, '<br>', "Deaths: ", deaths))
###########map############
##setup map
set_map_details <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
# Make sure both maps are on the same color scale
shadeLimit <- 9
# Create the map
fig<- plot_geo(cv_cfr,locationmode='USA-states') %>%
  add_trace(
    z= ~naive_CFR ,text= ~hover, location= ~state,
    color= ~naive_CFR, color='Purples'
  )


fig <- fig %>% colorbar(title="CFR May 1 2020", limits = c(0,shadeLimit))
fig <- fig %>% layout(
  title = paste('CFR by State as of', Sys.Date(), '<br>(Hover for value)'),
  geo = set_map_details
)

fig_May1 <- fig




### For Today

# Extract the data for each state by its abbreviation
cv_CFR <- cv_states_today %>%  select(state, abb, naive_CFR, cases, deaths) # select data
cv_CFR$state_name <- cv_CFR$state
cv_CFR$state <- cv_CFR$abb
cv_CFR$abb <- NULL

# Create hover text
cv_CFR$hover <- with(cv_CFR, paste(state_name, '<br>', "CFR: ", naive_CFR, '<br>', "Cases: ", cases, '<br>', "Deaths: ", deaths))

# Set up mapping details
set_map_details <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

# Create the map
fig <- plot_geo(cv_CFR, locationmode = 'USA-states') %>% 
  add_trace(
    z = ~naive_CFR, text = ~hover, locations = ~state,
    color = ~naive_CFR, colors = 'Purples'
  )
fig <- fig %>% colorbar(title = "CFR May 1 2020", limits = c(0,shadeLimit))
fig <- fig %>% layout(
    title = paste('CFR by State as of', Sys.Date(), '<br>(Hover for value)'),
    geo = set_map_details
  )
fig_Today <- fig


### Plot side by side 
### FINISH THE CODE HERE ###
subplot( fig_May1, fig_Today )

```





