---
title: "Lab 10 - SQL"
author: Tao Huang
output: tufte::tufte_html
link-citations: yes
---

# Setup

DELETE THIS CHUNK

```{r setup}
install.packages("RSQLite")
install.packages("DBI")


knitr::opts_chunk$set(eval = FALSE)
library(RSQLite)
library(DBI)
# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")
```



```{r}
# install.packages(c("RSQLite", "DBI"))
library(RSQLite)
library(DBI)
# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")
#con is temporary database which used to input data from other tables

# Download tables
actor <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/actor.csv")
rental <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/rental.csv")
customer <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/customer.csv")
payment <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/payment_p2007_01.csv")

# Copy data.frames to database
dbWriteTable(con, "actor", actor)
dbWriteTable(con, "rental", rental)
dbWriteTable(con, "customer", customer)
dbWriteTable(con, "payment", payment)

```

```{r}
dbListTables(con)
typeof(con)
str(actor)
```

TIP: Use can use the following QUERY to see the structure of a table

```{sql, connection=con}
-----#{sql} used to change to sql language.  #,connection="database" is neecessary

PRAGMA table_info(actor)          /*see the info of table "actor"*/

```

SQL references:

https://www.w3schools.com/sql/

# Exercise 1

Retrive the actor ID, first name and last name for all actors using the `actor` table. Sort by last name and then by first name.

```{sql, connection=con}

SELECT actor_id, first_name, last_name    
FROM actor
ORDER by last_name, first_name 

```

# Exercise 2

Retrive the actor ID, first name, and last name for actors whose last name equals 'WILLIAMS' or 'DAVIS'.

```{sql, connection=con}
SELECT actor_id, first_name, last_name
FROM actor
WHERE last_name IN ('WILLIAMS', 'DAVIS');    /*last_name=='WILLIAMS', 'DAVIS'*/
```

# Exercise 3

Write a query against the `rental` table that returns the IDs of the customers who rented a film on July 5, 2005 (use the rental.rental_date column, and you can use the date() function to ignore the time component). Include a single row for each distinct customer ID. 
```{sql, connection=con}
-----#{sql} used to change to sql language.  #,connection="database" is neecessary

PRAGMA table_info(rental)          /*see the info of table "actor"*/
```

```{sql, connection=con}
select date(rental_date)     /*data(..)  --extract date*/
from rental
```

```{sql, connection=con, output.var="July5th_ids"}

SELECT DISTINCT customer_id, rental_date  /*select distinct ---only return different value..ign*/
FROM rental
WHERE date(rental_date) = '2005-07-05'

```

```{r}
July5th_ids
```



# Exercise 4

## Exercise 4.1

Construct a query that retrives all rows from the `payment` table where the amount is either 1.99, 7.99, 9.99.

```{sql, connection=con}
/*see the info of table "actor"*/
pragma table_info(payment)
```

```{sql, connection=con}
SELECT *
FROM payment
WHERE amount IN (1.99, 7.99, 9.99)
```

## Exercise 4.2

Construct a query that retrives all rows from the `payment` table where the amount is greater then 5

```{sql, connection=con}
SELECT *
FROM payment
WHERE (amount >5)
```

## Exercise 4.2

Construct a query that retrives all rows from the `payment` table where the amount is greater then 5 and less then 8

```{sql, connection=con}
SELECT *
FROM payment
WHERE (amount>5) AND (amount<8)
```
# Exercise 5

Retrive all the payment IDs and their amount from the customers whose last name is 'DAVIS'.

```{sql, connection=con}
SELECT c.customer_id, p.amount, c.first_name, c.last_name
FROM customer AS c
  INNER JOIN payment AS p       /*join the two tables*/
ON c.customer_id=p.customer_id    /*customer_id is connection var.*/
where c.last_name = 'DAVIS'



```

# Exercise 6

## Exercise 6.1

Use `COUNT(*)` to count the number of rows in `rental`

```{sql, connection=con}

select count(*) AS n_obs
from rental

```

## Exercise 6.2

Use `COUNT(*)` and `GROUP BY` to count the number of rentals for each `customer_id`

```{sql, connection=con}
select count(*) AS n_obs2, customer_id
from rental
group by customer_id

```
## Exercise 6.3

Repeat the previous query and sort by the count in descending order

```{sql, connection=con}

select count(*) AS n_obs2, customer_id
from rental
group by customer_id
order by count(*) desc      /*descending order*/

```

## Exercise 6.4

Repeat the previous query but use `HAVING` to only keep the groups with 40 or more.

```{sql, connection=con}
select count(*) AS n_obs2, customer_id
from rental
group by customer_id
having count(*)>=40     /*only keep the groups with 40 or more.*/
order by count(*) desc      /*descending order*/

```

# Exercise 7

The following query calculates a number of summary statistics for the payment table using `MAX`, `MIN`, `AVG` and `SUM` of `amount`

```{sql, connection=con}
select max(amount) AS max, 
min(amount) AS min, 
avg(amount) AS avg,
sum(amount) AS sum

from payment

```

## Exercise 7.1

Modify the above query to do those calculations for each `customer_id`

```{sql, connection=con}
select max(amount) AS max, 
min(amount) AS min, 
avg(amount) AS avg,
sum(amount) AS sum,customer_id


from payment
group by customer_id
```
## Exercise 7.2

Modify the above query to only keep the `customer_id`s that have more then 5 payments

```{sql, connection=con}
select max(amount) AS max, 
min(amount) AS min, 
avg(amount) AS avg,
sum(amount) AS sum,customer_id,
count(*) AS count

from payment
group by customer_id
having count >5


```
# Cleanup

Run the following chunk to disconnect from the connection.

```{r}
# clean up
dbDisconnect(con)
```